---
title: "It's Seniority, Not Señoritas!"
author: "Maren Schubert"
date: "June 16, 2021"
output:
  html_document: 
    theme: spacelab
    toc: yes
    toc_float: yes
    smooth_scroll: yes
    fig_width: 10
    fig_height: 6
    fig_caption: yes
  pdf_document:
editor_options: 
  markdown: 
    wrap: 80
---

```{=html}
<style>
body {
text-align: justify}
</style>
```
```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = TRUE, # show results 
  collapse = TRUE, # not interrupt chunks 
  warning = FALSE, # suppress warning 
  message = FALSE, # suppress messages  
  tidy = 'styler', # style according to tidyverse guide
  cache = TRUE # cache the chunk outputs 
)

# packages needed 
p_needed <-
  c(
    "tidyverse", # for everything 
    "here", # for loading files 
    "broom", # for tidy model output 
    "skimr",
    "performance",
    "gridExtra",
    "modelsummary", # for model tables 
    "flextable", # for model tables 
    "MASS", # for mvrnorm
    "magrittr", # for pipes 
    "showtext", # for fonts 
    "checkdown", # for quizzes
    "viridis", # for colors
    "janitor", # for clean column names 
    "patchwork", # for ggplots side-by-side
    "haven", # for dta files 
    "latex2exp", # for math in graphs labels 
    "formatR", # for code formatting 
    "equatiomatic", # for model equations
    "separationplot", # for separation plots
    "lmtest", # for likelihood ratio test
    "tidymodels", # for nice ROC curves 
    "stargazer", # for model tables 
    "ggmosaic", # for mosaic plots
    "ggthemes", #for nice themes
    "mapproj", # for maps 
    "mfx", # for AME
    "maps", # for simple maps 
    "rgdal", # for shapefiles 
    "Zelig",
    "devtools",
    "hrbrthemes"
  ) 

devtools::install_github("IQSS/ZeligChoice")
library(ZeligChoice)

# check if packages are installed and if not, install them 
lapply(p_needed[!(p_needed %in% rownames(installed.packages()))], install.packages)

# prepare packages for use with library 
lapply(p_needed, library, character.only = TRUE) 

theme_set(
  theme_minimal() +
    theme(
      plot.title.position = "plot",
      legend.title = element_blank(),
      legend.position = "top"
    )
)

get_stats <- function(x) {
  if (is.double(x)){
  c(q = quantile(x, probs = c(0.005, 0.025, 0.05, 0.5, 0.95, 0.975, 0.995)), 
    mean = mean(x))}else{
      print("Non-numeric Data Passed")
    }
}
set.seed(12345)
```

**Many studies show a structural disadvantage of female legislators when it
comes to sponsoring bills and participating in the legislative process. But what
is the reason for the differences between men and women in parliament?**

Indeed, there are empirical observations, that female members of parliaments
sponsor significantly fewer bills than their male colleagues. A commonly cited
argument for those differences is the gender discrimination women are exposed
to. In short, other -- supposedly male - members of parliament may be
restricting women's access to political resources, reduce their effectiveness
and therefore cause fewer numbers of sponsored bills by women.

A study by Yann P. Kerevel and Lonna Rae Atkeson published in 2013 now suggests
that the reason of those smaller numbers of sponsored bills does not lie in the
gender of members of parliament, but in the time they spent in the legislative
institutions. Because in most parliaments there are many advantages for those
legislators who were in office for a long time. Women have only been taking up
public office in greater numbers relatively recently, and therefore had less
time to form networks, gain experience and get access to important resources and
posts in the process. This means, newer members of parliament -- who are very
often female -- will sponsor fewer bills than more experienced legislators.

This conclusion is drawn from the study with the title *Explaining the
Marginalization of Women in Legislative Institutions*. It is about bill
sponsoring processes in the Mexican Chamber of Deputies where there is no such
thing as incumbency as its members cannot be re-elected. Therefore, -- following
the authors of that study -- it is possible to examine exclusively the influence
of gender discrimination on bill sponsorship because there cannot be advantages
based on seniority.

As in many other states by now, there is a mandatory gender quota and although
some parties have tried to circumvent it, the proportion of women has increased
in the legislative periods following the law from 2000.

```{r data_preprocessing, include=FALSE}
replication <- read_dta(here("data", "JOPgender_replication_minversion.dta"))

bill_spon <- replication[c(1:42)]

bill_spon <- bill_spon %>%
  mutate(
    year = case_when(dum2000 == 1 ~ 2000,
                     dum2003 == 1 ~ 2003,
                     dum2006 == 1 ~ 2006
    )
  )
bill_spon <- bill_spon %>%
  filter(bills_presenter_total != "NA")

lt58 <- bill_spon %>%
  filter(year == "2000")

lt59 <- bill_spon %>%
  filter(year == "2003")

lt60 <- bill_spon %>%
  filter(year == "2006")


```

```{r, include=TRUE, echo=FALSE, fig.height=3}
cols <- c("0" = "#003056", "1" = "#df7e50")

na.omit(bill_spon) %>% 
  ggplot( 
       aes(x=year, 
           fill=factor(female))) + 
 geom_bar(
   position=position_dodge())+
  labs(
    title = "Number of Male and Female Legislators",
    subtitle = "In the 58th, 59th and 60th Legislative Term of the Mexican Chamber of Deputies"
  ) + 
  scale_color_manual(
    values = cols,
    aesthetics = c("fill"),
    labels = c("male", "female"),
    name = ""
  )  +
  theme_fivethirtyeight() +
  theme(
    axis.text.x = element_blank(),
    plot.subtitle = element_text(size=9),
    plot.title = element_text(size=18),
    legend.text = element_text(size=8)
    ) + 
  annotate(
    "text",
    label = "2000 - 2003",
    color = "#424242",
    family = "sans",
    x = 2000,
    y = -20,
    size = 3
  ) + 
  annotate(
    "text",
    label = "2003 - 2006",
    color = "#424242",
    family = "sans",
    x = 2003,
    y = -20,
    size = 3
  ) + 
  annotate(
    "text",
    label = "2006 - 2009",
 color = "#424242",
    family = "sans",
    x = 2006,
    y = -20,
    size = 3
  ) 
  
```

```{r, include=FALSE}
m1 <- glm.nb(bills_presenter_total ~ female + educ_level2 + p_diputadolocal + p_feddeputy+ p_senador + committee_chair + committee_secretary + party_leader + tier + prddummy + pandummy + pvemdummy + ptdummy + convdummy + minorparties_dum2 + wnom1_dist_chambermed + p_statepartyleader + p_natpartyleader, lt58 )

tidy(m1)

table_1 <- modelsummary(
  m1,  # including all models
  title = "Influence of Gender on Bill Sponsorhip in the Mexican Chamber of Deputies, 2000 (Replication)",
  notes = "Note: tba",
  output = "flextable",
  stars = TRUE, # including stars to show significance-levels,
  gof_omit = "IC|F|Log.", # excluding all matching gof statistics
  align = c("lcc"), # left, center, center ODER????
) %>% 
  autofit()
tidy(m1)

table_1
```

```{r include=FALSE}
m2 <- glm.nb(bills_presenter_total ~ female + educ_level2 + p_diputadolocal + p_feddeputy+ p_senador + committee_chair + committee_secretary + party_leader + tier + prddummy + pandummy + pvemdummy + ptdummy + convdummy + minorparties_dum2 + wnom1_dist_chambermed + p_statepartyleader + p_natpartyleader, lt59 )

tidy(m2)
table_2 <- modelsummary(
  m2,  # including all models
  title = "Influence of Gender on Bill Sponsorhip in the Mexican Chamber of Deputies, 2003 (Replication)",
  notes = "Note: tba",
  output = "flextable",
  stars = TRUE, # including stars to show significance-levels,
  gof_omit = "IC|F|Log.", # excluding all matching gof statistics
  align = c("lcc"), # left, center, center ODER????
) %>% 
  autofit()
tidy(m2)
table_2
```

```{r, include=FALSE}
m3 <- glm.nb(bills_presenter_total ~ female + educ_level2 + p_diputadolocal + p_feddeputy+ p_senador + committee_chair + committee_secretary + party_leader + tier + prddummy + pandummy + pvemdummy + ptdummy + convdummy + minorparties_dum2 + wnom1_dist_chambermed + p_statepartyleader + p_natpartyleader, lt60 )
tidy(m3)
table_3 <- modelsummary(
  m3,  # including all models
  title = "Influence of Gender on Bill Sponsorhip in the Mexican Chamber of Deputies, 2006 (Replication)",
  notes = "Note: tba",
  output = "flextable",
  stars = TRUE, # including stars to show significance-levels,
  gof_omit = "IC|F|Log.", # excluding all matching gof statistics
  align = c("lcc"), # left, center, center ODER????
) %>% 
  autofit()

table_3
```

```{r, include=FALSE, echo=FALSE}
## It's just for me and a tidier overview :)
models_sum <- list()
# specify how we want to name the models:
models_sum[["2000"]] <- m1
models_sum[["2003"]] <- m2
models_sum[["2006"]] <- m3

# models_sum[["Model 3"]] <- m3 ### fehlt noch

t_sum <- modelsummary(
  models_sum,  # including all models
  title = "Replication: Bill Sponsoring in Mexican Chamber of Deputies",
  notes = "Note: Standard errors are parentheses. ∗p < .05; ∗∗p < .01; ∗∗∗p < .001",
  output = "flextable", # defining the output we want
    stars = TRUE, # including stars to show significance-levels,
  gof_omit = "IC|F|Log.", # excluding all matching gof statistics
  align = c("lcc"), # left, center, center
) %>% 
  autofit()

t_sum
```

```{r, include=FALSE}
sim_betas_00 <- mvrnorm(1000, coef(m1), vcov(m1))
names(coef(m1))

medians_00 <- m1$model %>% ## take data used to estimate model
  as_tibble() %>% ## turn to tidy
  summarize_all(median)
```

```{r include=FALSE}
scen_m_00 <- cbind(1, # Intercept
                   medians_00[2:length(medians_00)]
)

scen_f_00 <- scen_m_00 %>%
  mutate(female = 1)

lambda_m_00 <- sim_betas_00 %*% t(scen_m_00) %>% 
  exp()

lambda_f_00 <- sim_betas_00 %*% t(scen_f_00) %>% ## 
  exp()

evs_m_00 <- apply(lambda_m_00, c(1, 2), function(x) {
  mean(rnegbin(1000, mu = x, theta = m1$theta)) #
})

evs_f_00 <- apply(lambda_f_00, c(1, 2), function(x) {
  mean(rnegbin(1000, mu = x, theta = m1$theta)) #
})
```

```{r, include=FALSE}
evs_m_00 %<>% as_tibble()
evs_f_00 %<>% as_tibble()
```

```{r, include=FALSE}
stats_m_00 <- get_stats(evs_m_00$V1)
stats_m_00 %<>% as_tibble

stats_f_00 <- get_stats(evs_f_00$V1)
stats_f_00 %<>% as_tibble
```

```{r include=FALSE}

p_lt_58 <- ggplot() + 
  geom_density(
    data = evs_m_00,
    mapping = aes(V1),
    color = "white",
    fill = "#003056",
    alpha = 0.8
  ) +
  geom_density(
    data = evs_f_00,
    mapping = aes(V1),
    color = "white",
    fill = "#df7e50",
    alpha = 0.8
  ) + 
  labs(
   title = "58th Legislative Term"
  ) + 
  geom_vline(
    aes(xintercept = 0.680975), lty = 2, color = "#003056"
    ) + # i had to set the lines by hand because the get_CI function didn't work
  geom_vline(
    aes(xintercept = 1.234150), lty = 2, color = "#003056"
    ) +
  geom_vline(
    aes(xintercept = 0.607975), lty = 2, color = "#df7e50"
    ) +
  geom_vline(
    aes(xintercept = 1.321050), lty = 2, color = "#df7e50"
    ) +
  theme_fivethirtyeight() +
  theme(
    plot.title = element_text(size = "10")
  ) 

p_lt_58 
```

```{r, include=FALSE}
sim_betas_03 <- mvrnorm(1000, coef(m2), vcov(m2))
names(coef(m2))

medians_03 <- m2$model %>% ## take data used to estimate model
  as_tibble() %>% ## turn to tidy
  summarize_all(median)
```

```{r, include=FALSE}
scen_m_03 <- cbind(1, # Intercept
                   medians_03[2:length(medians_03)]
)

scen_f_03 <- scen_m_03 %>%
  mutate(female = 1)

lambda_m_03 <- sim_betas_03 %*% t(scen_m_03) %>% 
  exp()

lambda_f_03 <- sim_betas_03 %*% t(scen_f_03) %>% ## 
  exp()

evs_m_03 <- apply(lambda_m_03, c(1, 2), function(x) {
  mean(rnegbin(1000, mu = x, theta = m2$theta)) #
})

evs_f_03 <- apply(lambda_f_03, c(1, 2), function(x) {
  mean(rnegbin(1000, mu = x, theta = m2$theta)) #
})
```

```{r, include=FALSE}
evs_m_03 %<>% as_tibble()
evs_f_03 %<>% as_tibble()
```

```{r, include=FALSE}
stats_m_03 <- get_stats(evs_m_03$V1)
stats_m_03 %<>% as_tibble

stats_f_03 <- get_stats(evs_f_03$V1)
stats_f_03 %<>% as_tibble
```

```{r, include=FALSE}
p_lt_59 <- ggplot() + 
  geom_density(
    data = evs_m_03,
    mapping = aes(V1),
    color = "white",
    fill = "#003056",
    alpha = 0.8
  ) +
  geom_density(
    data = evs_f_03,
    mapping = aes(V1),
    color = "white",
    fill = "#df7e50",
    alpha = 0.8
  ) + 
  labs(
    title= "59th Legislative Term"
  ) + geom_vline(
    aes(xintercept =2.459975), lty = 2, color = "#003056"
    ) +
  geom_vline(
    aes(xintercept = 4.083025), lty = 2, color = "#003056"
    ) +
  geom_vline(
    aes(xintercept = 2.653900), lty = 2, color = "#df7e50"
    ) +
  geom_vline(
    aes(xintercept = 5.014425), lty = 2, color = "#df7e50"
    ) +
  theme_fivethirtyeight() +
  theme(
    plot.title = element_text(size = "10")
  )
p_lt_59
```

```{r, include=FALSE}
sim_betas_06 <- mvrnorm(1000, coef(m3), vcov(m3))
names(coef(m3))

medians_06 <- m3$model %>% ## take data used to estimate model
  as_tibble() %>% ## turn to tidy
  summarize_all(median)
```

```{r, include=FALSE}
scen_m_06 <- cbind(1, # Intercept
                   medians_06[2:length(medians_06)]
)

scen_f_06 <- scen_m_06 %>%
  mutate(female = 1)

lambda_m_06 <- sim_betas_06 %*% t(scen_m_06) %>% 
  exp()

lambda_f_06 <- sim_betas_06 %*% t(scen_f_06) %>% ## 
  exp()

evs_m_06 <- apply(lambda_m_06, c(1, 2), function(x) {
  mean(rnegbin(1000, mu = x, theta = m3$theta)) #
})

evs_f_06 <- apply(lambda_f_06, c(1, 2), function(x) {
  mean(rnegbin(1000, mu = x, theta = m3$theta)) #
})
```

```{r, include=FALSE}
evs_m_06 %<>% as_tibble()
evs_f_06 %<>% as_tibble()
```

```{r, include=FALSE}
stats_m_06 <- get_stats(evs_m_06$V1)
stats_m_06 %<>% as_tibble

stats_f_06 <- get_stats(evs_f_06$V1)
stats_f_06 %<>% as_tibble
```

```{r, include=FALSE, echo=FALSE}
p_lt_60 <- ggplot() + 
  geom_density(
    data = evs_m_06,
    mapping = aes(V1),
    color = "white",
    fill = "#003056",
    alpha = 0.8
  ) +
  geom_density(
    data = evs_f_06,
    mapping = aes(V1),
    color = "white",
    fill = "#df7e50",
    alpha = 0.8
  ) + 
  labs(
    title= "60th Legislative Term"
  ) + geom_vline(
    aes(xintercept = 6.533000), lty = 2, color = "#003056"
    ) +
  geom_vline(
    aes(xintercept = 11.923775), lty = 2, color = "#003056"
    ) +
  geom_vline(
    aes(xintercept = 6.538775), lty = 2, color = "#df7e50"
    ) +
  geom_vline(
    aes(xintercept = 12.862525), lty = 2, color = "#df7e50"
    ) +
  theme_fivethirtyeight() +
  theme(
    plot.title = element_text(size = "10")
  )
p_lt_60
```

Apart from the distribution of female and male representatives, the numbers of
bills sponsored by the members of those two groups are the values of interest.
The plots below show the numbers of bills sponsored by an average male or female
legislator in the three legislative terms considered. The mostly overlapping
curves show that on average there is no significant difference between male and
female members of the Mexican Chamber of Deputies.

```{r include=TRUE, echo=FALSE, fig.height=3}
p_bind <- p_lt_58 + p_lt_59 + p_lt_60

p_bind + plot_annotation(
  title = 'Number of Bills Sponsored \nin the Mexican Chamber of Deputies',
  subtitle = 'By Average Female and Male Representatives', 
  theme = theme_fivethirtyeight())
```

Interestingly, in the first legislative term analysed an average man or woman
only sponsored one bill. In the 59th legislative term the representatives were
more active: the average man sponsored 3 bills, ranging between 2.5 and 4 bills.
At the same time the average woman even sponsored 3.7 bills, plus or minus 1.5
bills. That shows, that in the first term after the effective mandatory gender
quota was implemented, women sponsored slightly more bills, but still the
difference is not significant. Lastly, in the 60th legislative term the curves
are mainly overlapping. An average woman sponsored 9.5 bills, plus or minus 3,
while an average man sponsored 9 bills, plus or minus 2.5. Like in the 59th
term, one cannot speak of significant differences between men and women in terms
of bill sponsorship.

Based on their results the authors come to the conclusion that reasons for
differences in bill sponsorship of female and male legislators in other studies
can mainly be explained by seniority and not by the gender of the members of
parliament.
